### 2024-09-10 202409_aij_dc_problog_aij_second/files_old/preliminaries.tex

\section{Preliminaries}

We expect the reader to be familiar with basic language concepts and terms of Prolog \todo{cite (Lloyd, 1989; Flach, 1994)}; a quick summary can be found in~\ref{app:lp}. \ak{this sentence and appendix copied from plp concepts paper; should go earlier given we already have examples there?}

\subsection{Probabilistic Logic Programming}
\ak{how much do we need here, given that we later introduce ProbLog constructs through their mappings to the new language? how much detail on the distribution semantics is needed for completeness? (the semantics sec is heavily linked to the Sato paper, but we can't repeat all of that)}

\subsection{Weighted Model Integration}

\todo{introduce all the wmc, amc, wmi gadgets}